---
title: 'ESM 204 #4'
author: "Sage Kime and Karla Garibay Garcia"
date: "5/17/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(stargazer)
library(broom)
options(scipen = 0)
```


### 1. Quadratic Damage Function 
Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Omit an intercept term; damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r, include=FALSE, warning=FALSE, message=F}
# Read in data
warming <- read.csv("warming.csv")
damages <- read.csv("damages.csv")

#create new variable for warming^2 (followed tutorial from statology.org/quadratic-regression-r/)
damages$warming2 <- damages$warming^2

#fit quadratic regression model
quadmodel <- lm(damages ~ warming + warming2 +0, data=damages)

summary(quadmodel)

# damages = -1.035e12x + 1.94e13x^2

warming_values <- seq(0,10,0.1)

#create list of predicted warming levels using quadratic model
predicted_damages <- predict(quadmodel, list(warming=warming_values, warming2=warming_values^2))

#plot with quad regression 
plot(damages$warming, damages$damages) +
  lines(warming_values, predicted_damages, col='blue')

# From Ian
damages_model <- function(x) {
  quadmodel$coefficients[[1]]*x + quadmodel$coefficients[[2]]*(x^2)
}

```

```{r, echo=FALSE, message=F}
#ggplot
ggplot(data = damages, aes(x = warming, y = damages)) +
  geom_point(alpha = .5,
              size = 2,
              color = "darkred",
              ) +
    labs(y = "Financial Damages ($)", title = "Estimated annual total damages from climate change based on warming", x = "Degrees of Warming (C)") +
    theme_minimal() +
  stat_function(fun = damages_model, size = 1)

```

### 2. Damages in each year under the baseline climate and pulse scenarios
Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

```{r, include=FALSE, message=F}
#Damages WITHOUT pulse

# Name coefficients
a <- quadmodel$coefficients[[1]]
b <- quadmodel$coefficients[[2]]

#create function
base_damage <- function(x) {
  baseline_damage <- quadmodel$coefficients[[1]]*x + quadmodel$coefficients[[2]]*(x^2)
  return(baseline_damage)
}

#Ian
damages_model <- function(x) {
  quadmodel$coefficients[[1]]*x + quadmodel$coefficients[[2]]*(x^2)
}

#test fuction - create vector
base_damage(warming$warming_baseline)

#convert vector to variable in warming dataset
warming$damages_baseline <- base_damage(warming$warming_baseline)


#Damages WITH pulse

#create function
pul_damage <- function(x) {
  pulse_damage <- quadmodel$coefficients[[1]]*x + quadmodel$coefficients[[2]]*x^2
  return(pulse_damage)
}

#test fuction - create vector
pul_damage(warming$warming_pulse)

#convert vector to variable in warming dataset
warming$damages_pulse <- pul_damage(warming$warming_pulse)


#Difference between pulse and basline
warming$damages_diff <- warming$damages_pulse - warming$damages_baseline

#Difference per ton
warming$damages_ton <- warming$damages_diff/35e9


plot(warming$year, warming$damages_baseline) #DO ALL PLOTS IN GGPLOT
plot(warming$year, warming$damages_pulse)
plot(warming$year, warming$damages_diff)
plot(warming$year, warming$damages_ton)
```

```{r, echo=F}
#ggplot - baseline
ggplot(data = warming, aes(x = year, y = damages_baseline)) +
  geom_point(alpha = .5,
              size = 2,
              color = "blue",
              ) +
    labs(y = "Financial Damages ($)", title = "Estimated annual total damages from climate change in baseline scenario", x = "Year") +
    theme_minimal()

#ggplot - pulse
ggplot(data = warming, aes(x = year, y = damages_pulse)) +
  geom_point(alpha = .5,
              size = 2,
              color = "blue",
              ) +
    labs(y = "Financial Damages ($)", title = "Estimated annual total damages from climate change in pulse scenario", x = "Year") +
    theme_minimal()

#ggplot - difference
ggplot(data = warming, aes(x = year, y = damages_diff)) +
  geom_point(alpha = 1,
              size = 1,
              color = "darkorange",
              ) +
    labs(y = "Financial Damages ($)", title = "Differences in damages over time between baseline and pulse scenarios", x = "Year") +
    theme_minimal()

#ggplot - per ton
ggplot(data = warming, aes(x = year, y = damages_ton)) +
  geom_point(alpha = 1,
              size = 1,
              color = "blue",
              ) +
    labs(y = "Financial Damages ($)", title = "Difference in damages over time from the pulse per ton of
CO2", x = "Year") +
    theme_minimal()
```

### 3. Calculate the SCC
The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates.

```{r, echo = F}
#SCC = PV(pulse damages) - PV(baseline damages)
#PV = Xt / (1+r)^t

# Calculate PV using ~for loop~
PV_calc <- function(values, discount_rate) {
  sum <- 0;
  r <- discount_rate
  for (i in c(1:length(values))) {
    current <- values[i]/(1 + r)^i
    sum <- sum + current
  }
  return(sum)
}

# Defining "reasonable" range of discount rate
discount_rates <- seq(.01,.1,.005)

# Calculate SCC values at determined discount rates
SCC_values <- PV_calc(warming$damages_ton, discount_rates)

# Store as data frame
SCC_data <- data.frame(discount_rates, SCC_values)

# ggplot of SCC values at various discount rates
ggplot(data = SCC_data, aes(x = discount_rates, y = SCC_values)) +
  geom_line(alpha = .9,
              size = 1,
              color = "darkgray",
              ) +
    labs(y = "SCC Value ($)", title = "Social Cost of Carbon (SCC) per ton of CO2 at various discount rates", x = "Discount Rate (0-10%)") +
    theme_minimal() +
  annotate(geom = "point", x=0.021, y=70.65026, shape = "circle", color = "darkred", size = 3) +
    annotate("text", x=0.035, y=70.65026, label = "SCC w/ Ramsey's Rule", size = 3.5) +
  geom_segment(aes(x=0, xend=.021, y=70.65026, yend=70.65026), colour="darkred", linetype = "dashed") +
    geom_segment(aes(x=.021, xend=.021, y=0, yend=70.65026), colour="darkred", linetype = "dashed")

```


### 4. Ramsey's Rule
The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation: r = ρ + ηg
Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

```{r}
#Ramsey's Rule: discount rate
r <- .001 + 2*.01

#SCC at discount rate of .021 (Ramsey's Rule)
PV_calc(warming$damages_ton, .021)
```

### 5. Policy A vs. Policy B

What is the expected present value of damages up to 2100 under Policy A? What is the
expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy
A costs zero and undertaking Policy B costs X. How large could X be for it to still make
economic sense to pursue Policy B instead of Policy A? Qualitatively, how would your answer
change if society were risk averse?

```{r}
#Policy A

#Calculate PV using ~for loop~
PV_calc <- function(values, discount_rate) {
  sum <- 0;
  r <- discount_rate
  for (i in c(1:length(values))) {
    current <- values[i]/(1 + r)^i
    sum <- sum + current
  }
  return(sum)
}

#Create new variable for 1.5x baseline in policy A
warming$baseline150 <- warming$warming_baseline*1.5

#Create new variable for damages under 1.5 baseline outcome
warming$damages_baseline150 <- damages_model(warming$baseline150)

#Calculate PV of each outcome in policy A
A_baseline <- PV_calc(warming$damages_baseline, 0.02)
A_baseline150 <- PV_calc(warming$damages_baseline150, 0.02)

expected_value <- .5*A_baseline + .5*A_baseline150
expected_value


#Policy B

#Create new variable for warming until 2050
warming <- warming %>% 
  mutate(warming_baseline_2050 = 
           case_when(year > 2050 ~ 1.29,
                     TRUE ~ warming_baseline
          ))

#Create new variable for damages under 2050 baseline outcome
warming$damages_baseline2050 <- damages_model(warming$warming_baseline_2050)

#Calculate PV of policy B
B_baseline <- PV_calc(warming$damages_baseline2050, 0.02) #PV is expected value bc only one option


#Subtract cost of policy A - policy B (SOS - CHECK THIS)
max_x <- expected_value - B_baseline
max_x

```

